---
title: "Qualitative Analysis of Reddit Data Using R"

output: html_notebook
bibliography: dissertation_bibliography.bib
csl: apa.csl
---
### Gentry Hanks, Queen's University
# Abstract
@adams_boundless_2005
The embodiment of surveillant technologies provides a means and site of production of data to be consumed by both the self and the medical gaze with physical and emotional consequences. Analyzing discourse on the subreddit r/diabetes, I examine assemblages of surveillant technologies that render an ever-increasing quantified self for those using insulin pumps and glucose monitors. Haraway (1990) brought cyborgs to the fore in the early 1990s and Lupton (2000) describes cyborgs ``When hooked up to medical (and other) technologies, the patient's body becomes a cyborg, a juncture of human flesh and machine'' (p.56). Bodies are rendered regulatable through the use of embodied and disembodied technologies. People with any type of diabetes may be treated with insulin, which can be self-administered through multiple daily injections or through an insulin pump. Insulin pumps, as well as insulin, are proprietary. There is a growing do-it-yourself movement when it comes to hacking the cyborg self. Open source communities have made headway in generating new technology, reappropriating old devices or implementing everyday hacks of hardware and its fleshy interface (Forlano, 2016). Lupton (2016) describes these devices and data as intermingling within a data economy, which I argue in the case of diabetes are used in surveillance and the medical gaze. Devices used to manage diabetes quantify the self and datify the device user. These data as perceived by medical practitioners, family members, friends, strangers and last, but not least, the self can have significant effects on everyday life, socio-spatial relations and emotional health.


```{r}
library(dplyr) # dplyr is library written by Hadley Wickham for easy manipulation of data. See https://cran.r-project.org/web/packages/dplyr/index.html to download or http://dplyr.tidyverse.org/
library(readr)
library(ggplot2) # This is a plotting system for R, also by Hadley Wickham. See http://ggplot2.org/.
```
```{r, message=FALSE}
#a data frame is a data structure like a spreadsheet that uses code instead of a graphical user interface.


d <- read_csv("~/Downloads/reddit_diabetes.csv") #This data set was accessed through the reddit API and is publicly available.
```

```{r}
d %>%
  mutate(label = ifelse(grepl("pump", author_flair_text, ignore.case=TRUE) &
                          grepl("MDI", author_flair_text, ignore.case=TRUE), "pump & MDI", 
                        ifelse(grepl("pump", author_flair_text), "pumper",
                               ifelse(grepl("MDI", author_flair_text, ignore.case=TRUE), "MDI", NA)))) %>%
  select(author, author_flair_text, label) %>%
  filter(label == "pump & MDI" & !duplicated(author)) %>%
  as.data.frame() %>%
  nrow()


#Based on flair, how many authors have changed their flair from MDI to pump or pump to MDI?


d %>%
  mutate(label = ifelse(grepl("pump", author_flair_text, ignore.case=TRUE) &
                          grepl("MDI", author_flair_text, ignore.case=TRUE), "pump & MDI", 
                        ifelse(grepl("pump", author_flair_text), "pumper",
                               ifelse(grepl("MDI", author_flair_text, ignore.case=TRUE), "MDI", NA)))) %>%
  select(author, author_flair_text, label, created_utc) %>%
  # mutate(created_utc = as.POSIXct(created_utc, origin = "1970-01-01")) %>%
  # arrange(author, created_utc)
  filter(!is.na(label)) %>%
  group_by(author) %>%
  filter(!duplicated(label)) %>%
  count(author, sort = TRUE)

```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file).
